---
title: "Vaccinated in ICUs - Greece"
author: "Kostas Voulgaropoulos"
date: "12/2/2022"
output: html_document
---

## Purpose

The great bet with COVID is to avoid 
* overcrowding of Intensive Care Units (ICUs) and  
* exhaustion of medical personnel

Most scientists say that our **most powerful weapon** towards that cause is the **vaccine**.

Thankfully for us, the Greek Health Administration started (since August the 30th) publishing the numbers of vaccinated and unvaccinated patients in the country's Intensive Care Units.

We will donwload all available daily reports that contain those figures to investigate who's occupying the ICUs. 

Vaxed, unvaxed or both?

So, after fetching all available data, our dataset looks like:

```{r find data, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

#Libraries#######
library(tidyverse)
library(pdftools)
library(ggthemes)
library(lubridate)

Sys.setlocale("LC_ALL","Greek")


#Global Variables######


#Whether we will download the PDFs containing info
#TRUE for first run, FALSE if you have already collected and stored wanted data
for_download <- TRUE

#Whether we will ask for a fraction of the available PDFs
on_development <- FALSE

#Whether we want to leave plots as PNGs in working directory
leave_trail <- FALSE

#in case leave_trail is TRUE, the dimensions of the PNGs exported
wanted_width <- 1300
wanted_height <- 900

#PDF Links construction#####
#Create a character vector of the links where the pdfs are
today <- Sys.Date() 

#30th of August is the first day when numbers of (un)vaccinated patients reported
first_report <- as.Date.character("2021-08-30")

#Pattern that will return us numbers
patt_vaxed <- "[\\d\\.\\(%\\)\\.]+ .* ανεμβολίαστοι"

#Where we're gonna store our downloaded data
saved_data <- "EOPY.RData"

#Download or load from previous runs######
if(for_download){
  
  #Check whether we have data from previous runs
  data_already_exist <- file.exists(saved_data)
  
  #If we have
  if(data_already_exist){
    
    #Load the saved dataframe
    load(saved_data)
    
    #Store it to a differently named dataframe
    already_downloaded_data <- tem_df
    
    last_date_of_downloaded_data <- already_downloaded_data$Day %>% max() 
    
  }

  #All days between first day of report and today
  dates_wanted <- 
    #If it's run later than 15:00, include current day
    if(lubridate::hour(Sys.time()) > 15){
      
      seq.Date(from = first_report, to = today, by = 1)
      
    }else{
      seq.Date(from = first_report, to = today-1, by = 1)
      }

  if(data_already_exist){
    
    dates_wanted <- dates_wanted[dates_wanted > last_date_of_downloaded_data]
    
  }

  if(on_development){
    
    #Pick some random dates to minimize the iterations
    dates_wanted <- dates_wanted[sort(sample(x = 1:length(dates_wanted), size = 3, replace = FALSE))]
    
  }

  #In the form that each date appears in each link
  days_wanted <- 
    dates_wanted %>% 
    lubridate::month() %>% 
    as.character() %>% 
    str_pad(2,"left","0") %>%
    paste0(dates_wanted %>% lubridate::day() %>% as.character() %>% str_pad(2,"left","0"))
  
  #for some reason, 22/09 is represented as 09222 in its link
  days_wanted[days_wanted == "0922"] <- "09222" 
  
  #Finally create links
  links <- paste0("https://eody.gov.gr/wp-content/uploads/", 
                  lubridate::year(dates_wanted) ,
                  "/",
                  substr(days_wanted,1,2),
                  "/covid-gr-daily-report-", 
                  lubridate::year(dates_wanted),
                  days_wanted,
                  ".pdf")  


  
  
  
  #initiate a dataframe
  tem_df <- data.frame()
  
  #To Store unsuccessful attempts
  not_existing_links <- character()
  unsuccessful_links <- character()
  
  r <- 1
  for(j in 1:length(links)){
    
    print(paste("Day", dates_wanted[j]))
    
    link <- links[j]
    
    data <- tryCatch(pdftools::pdf_text(link), error = function(e) e)
    
    #If for some reasong pdf data does not get downlaoded
    if(any(class(data) == "error")){
      
      #append the vector where we store them
      not_existing_links <- c( not_existing_links, link)
      #And on to the next link
      next
      
    }
    
    #Locate where our pattern exists
    where_wanted_text <- data %>% sapply(str_locate,patt_vaxed,USE.NAMES = FALSE)
    
    #Specify the page
    which_page <- which(!is.na(where_wanted_text[1,]))
    
    
    if(identical(which_page,integer(0))){
      
      print(paste("No data found for ", link))
      
      #Store it to examine it later
      unsuccessful_links <- c(unsuccessful_links, link)
      
      next
      
    }else{
      
      tem_df[r,"File"] <- days_wanted[r]
      
      tem_df[r,"Day"] <- dates_wanted[r]
      
      lower_border <- where_wanted_text[,which_page][1] 
      
      upper_border <- where_wanted_text[,which_page][2]
      
      tem <- data[which_page] %>% 
        substr(lower_border-90, upper_border+90) %>% 
        str_extract_all("\\d{2,} \\(\\d[\\d\\.]+%\\)")
      
      tem_df[r,"Unvaccinated"] <- tem[[1]][1]  
      
      tem_df[r,"Vaccinated"] <- tem[[1]][2]  
      
      tem_df[r,"link"] <- link
      
      r <- r+1
      
    }
    
  }
  
  
  
  #Bind with data from previous session(s)
  if(data_already_exist){tem_df <- rbind( already_downloaded_data,tem_df) }
  
  #And save for next sessions
  save(tem_df, file = saved_data)
  
  
}else{
  
  load(saved_data)

}


#Tidy up Data############
#Separate absolute numbers from percentages
for(col in c("Unvaccinated", "Vaccinated")){
  
  tem_df <- 
    tem_df %>% separate( col = col, into = c(col,paste0(col,"_perc")), sep = "\\s" ) #%>% pull(col)
  
}

#Clean Percentages
for(col in paste0(c("Unvaccinated", "Vaccinated"),"_perc")){
  
  tem_df[,col] <-
    tem_df[,col] %>% str_extract("\\d+\\.\\d+") %>% as.numeric()

}


tem_df %>% head()


```


## Absolute Numbers & Percentages

Now that we have fetched and cleaned our data, we can plot.

First, let's take a look at the absolute numbers in ICUs.


```{r absolute numbers, echo=FALSE}

#Now change the locale so that plots can be read in English too
invisible(Sys.setlocale("LC_ALL","English"))


unvaccinated_color <- "#a78ccf" #"#ed8282" #"#e07e7e"
vaccinated_color <- "#69c28b"


if(leave_trail){
  png(filename="Absolute_Numbers_in_ICUs.png", 
      width=wanted_width, 
      height=wanted_height)
}



in_long_format <- 
  
  tem_df %>% 
  
  pivot_longer(cols = c("Unvaccinated", "Vaccinated")) %>% 
  
  mutate(value = as.integer(value), FillColor = ifelse(name == "Unvaccinated", unvaccinated_color , vaccinated_color ) ) 
    

maximum_number_of_vaccinated <- 
  
  in_long_format %>% filter(name=="Vaccinated") %>%
  
  pull(value) %>% max()


day_of_maximum_number_of_vaccinated <- 
  
  in_long_format %>% filter(value==maximum_number_of_vaccinated) %>%
  
  pull(Day) 


absolutes_plot <- in_long_format %>% 
  
  ggplot(aes(x = Day, y = value, fill = name)) +
  
  geom_col(position = "stack",width = .5) + 
  
  scale_fill_manual(values = in_long_format$FillColor) +

  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  
  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%d %b %y") +
  
  scale_y_continuous(breaks = sort(c(maximum_number_of_vaccinated,seq(100, 800, 100)) )) +
  
  geom_hline(yintercept = maximum_number_of_vaccinated, color = "red") +

  theme_fivethirtyeight() + 
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1) , legend.title = element_blank() ) +
          
  labs(title = "Number of COVID Patients", subtitle =  "In Intensive Care Units in Greece", y = "Number of Patients", x = "") + 
  
  geom_label(x = day_of_maximum_number_of_vaccinated, 
             
             y = maximum_number_of_vaccinated+60, 
             
             label = maximum_number_of_vaccinated,show.legend = FALSE)


if(leave_trail){dev.off()}else{absolutes_plot}



```

By naked eye, it appears that around 2/3 of the total patients are unvaccinated.

And one with Relative.

```{r relative numbers }

#And one with percentages

if(leave_trail){
  png(filename="Percentages_in_ICUs.png", 
      width=wanted_width, 
      height=wanted_height)
}

daily_numbers <- in_long_format %>% group_by(Day) %>% summarize(Daily_Total = sum(value))


percentages_df <- 
  
  in_long_format %>% 
  
  left_join(daily_numbers, by = "Day") %>% 
  
  mutate( Percentage = value / Daily_Total ) 


maximum_percentage_of_vaccinated <- 
  
  percentages_df %>%
  
  filter(name == "Vaccinated") %>%
  
  pull(Percentage) %>% max()
  
day_of_maximum_percentage <- 
  
  percentages_df %>% 
  
  filter(Percentage == maximum_percentage_of_vaccinated) %>%
  
  pull(Day)

maximum_percentage_of_vaccinated <- round(maximum_percentage_of_vaccinated*100)


percentage_breaks <- seq(0,100,25)

percentage_breaks <- sort(c(percentage_breaks))


percentages_plot <- 
  
  percentages_df %>%
  
  ggplot(aes(x = Day, y = Percentage*100, fill = name )) +
  
  geom_col(position = "stack",width = .5) + 
  
  scale_fill_manual(values = c(unvaccinated_color , vaccinated_color) ) +

  geom_hline(yintercept = maximum_percentage_of_vaccinated, color = "red") +

  theme_fivethirtyeight() + 
    
  theme(axis.text.x = element_blank(), 
        legend.title = element_blank()) +
  
  scale_y_continuous( breaks = percentage_breaks ,labels = paste0(percentage_breaks,"%") ) +

  scale_x_date(date_breaks = "2 weeks", date_minor_breaks = "1 week", date_labels = "%d %b %y") +
  
  labs(title = "Percentages of COVID Patients", subtitle = "In Intensive Care Units in Greece", y = "%", x = "") + 
  
  geom_label(x = day_of_maximum_percentage, 
             
             y = maximum_percentage_of_vaccinated+10, 
             
             label = paste0(maximum_percentage_of_vaccinated,"%"), show.legend = FALSE )


if(leave_trail){dev.off()}else{percentages_plot}

```

## Distribution of Populations



By taking a look at the distributions of the numbers exhibited each day, we examine their numbers regardless of the days they occurred.


```{r distribution}


#Distribution
if(leave_trail){
png(filename="distribution.png", width=wanted_width+50, height=wanted_height-50)
}

tem_df %>% 
  
  pivot_longer(cols = c("Unvaccinated", "Vaccinated")) %>% 
  
  mutate(value = as.integer(value)) %>% 
  
  ggplot(aes(x = value, fill = name )) +
  
  geom_density(alpha = 0.7) + 
  
  theme_fivethirtyeight() + 
  
  labs(title = "Distribution of Daily Number of Patients per Group", 
       x = "Daily Count of Patients", 
       y = "Frequency") + 

  scale_fill_manual(values = c(unvaccinated_color , vaccinated_color) )
  
if(leave_trail){dev.off()}

```

The numbers "scored" by the two groups make their distinction so blatant.

The highest numbers recorded by the vaccinated are even lower than the lowest of unvaccinated.


# Conlusion

Even though this current report is not a piece of extensive research and cannot directly suggest probabilities of ICU hospitalization, it provides a strong indication that vaccination contributes in avoiding the congestion of Intensive Care Units.